Automated iterative near-term forecasting for the Portal Project

### 2018-02-07 manuscript/portal_iterative_forecasting_ms.Rmd
Manuscript polishing
1. Reduce redundancy in abstract
2. General wording and tightening throughout

---
title: "Automated iterative near-term forecasting for the Portal Project"
csl: mee.csl
fontsize: 12pt
output:
  html_document: default
  pdf_document:
    keep_tex: yes
header-includes:
- \usepackage{times}
- \usepackage{setspace}
- \usepackage{booktabs}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
bibliography: refs.bib
---
\raggedright

@@

Forecasting the future state of ecological systems is important for management, conservation, and evaluation of our basic understanding of how ecology operates [@clark2001; @tallis2006; @diaz2015; @dietze2017]. In 2001, Clark et al. [-@clark2001] called for a more central role of forecasting in ecology, and since then an increasing number of ecological forecasts are being published. However, most of these forecasts are made once, published, and never assessed or updated. Without assessment, we have limited information on how much confidence to place in our predictions; without regular updates, forecasts lack the most up-to-date information as conditions change [@dietze2018]. This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions. More regular updating and assessment will advance ecological forecasting as a field by speeding up the identification of the best models for forecasting specific issues in specific ecosystems and contribute overall to our understanding of how to best design forecasting approaches for ecology. For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].

With the goal of making ecological forecasting more dynamic and responsive, Dietze et al [-@dietze2018] recently called for an increase in iterative near-term forecasting. Iterative near-term forecasting means making forecasts for the near future and making these forecasts repeatedly through a cycle of forecast evaluation, integration of updated data, and generation of new forecasts. This approach to forecasting has a number of advantages. Because forecasts are made 'near-term'---daily to annual forecasting instead of multi-decadal---predictions can be assessed more quickly and frequently, leading to more rapid model improvements [@tredennick2016; @dietze2018]. Because the forecasts are made repeatedly through time, new data can be integrated with each new iteration of the forecast cycle. This iterative approach to forecasting allows any changes in the state of the system that have occurred since the previous forecast to be incorporated and accounted for, thereby improving the accuracy of future forecasts [@dietze2018]. Iterative near-term forecasting has the potential to promote rapid improvement in the state of ecological forecasting by quickly identifying how models are failing, facilitating rapid testing of improved models, and incorporating updated data so models run with the most up-to-date information on the system available. While use of iterative near-term forecasting is often discussed in the context of policy and management [@clark2001; @luo2011; @petchey2015], this approach to model testing can also be used to improve our basic understanding of ecological systems [@dietze2017]. For example, alternative mechanistic models can be competed against each other to see which model provides the best forecasts for near-term dynamics, thus providing insights into the relative importance of different processes driving dynamics of ecological systems [@dietze2018]. Whether deployed for basic or applied uses, iterative near-term forecasting incorporates a more dynamic interplay between models, predictions, and data that is clearly needed to improve ecological forecasting and our understanding of ecological systems more broadly.

Because iterative near-term forecasting requires a dynamic interplay of models, predictions, and data, Dietze et al [-@dietze2018] highlight approaches to data management, model construction and evaluation, and cyberinfrastructure that are necessary to effectively implement this type of forecasting (Box 1). Data to be used for iterative near-term forecasting needs to be widely accessible, which requires data to be released quickly under open licenses [@dietze2018; @vargas2017] and structured so that it can be used easily by a variety of researchers and in multiple modeling approaches [@borer2009; @strasser2011]. Models need to be able to deal with uncertainty, in the predictors and the predictions, to properly convey uncertainty in the resulting forecasts [@diniz-filho2009]. Multiple models should be developed, both to assess which models are performing best [@dietze2018] and to facilitate combining models to form ensemble predictions which tend to perform better than single models [@araujo2007; @diniz-filho2009]. Ensuring that data and models are regularly updated and new forecasts are made requires cyberinfrastructure to automate data processing, model fitting, prediction, model evaluation, forecast visualization, and archiving. In combination, these approaches should allow forecasts to be easily rerun and evaluated as new data becomes available [Box 1; @dietze2018].

While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement; few of their recommendations are in widespread use in ecology today. We examined what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going long-term (~40 year) ecological study that collects high-frequency data on desert rodent abundances [@brown1998; @ernest2008]. We constructed our forecasting pipeline with the goal of being able to forecast rodent abundances and evaluate our predictions on a monthly basis. In this paper, we discuss our approach for creating this iterative near-term forecasting pipeline, the challenges we encountered, the tools we used, and the lessons we learned so that others can create their own iterative forecasting systems.

@@

Iterative forecasting requires data that is collected repeatedly, and it benefits most from data that are collected frequently, as this provides more opportunities for updating model results and assessing (and potentially improving) model performance [Box 1; @dietze2018]. The Portal Project is a long-term ecological study situated in the Chihuahuan Desert (2 km north and 6.5 km east of Portal, Arizona, US). Researchers have been continuously collecting data at the site since 1977, including data on the abundance of rodent and plant species (monthly and twice yearly, respectively) and climatic factors such as air temperature and precipitation (daily) [@brown1998; @ernest2009; @ernest2016]. The site consists of 24 50m x 50m experimental plots. Each plot contains 49 permanently marked trapping stations laid out in a 7 x 7 grid, and all plots are trapped with Sherman live traps for one night each month. For all rodents caught during a trapping session, information on species identity, size, and reproductive condition is collected, and new individuals are given identification tags. This information on rodent populations is high-frequency, uses consistent trapping methodology, and has an extended time-series (469 monthly samples and counting), making this study an ideal case for near-term iterative forecasting.

@@

Implementation of iterative forecasting requires the regular rebuilding of models with new raw data as it becomes available and the presentation of those forecasts in usable forms; in our case, this occurs monthly. Rebuilding models in an efficient and maintainable way relies on developing an automated pipeline to handle the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presenting of the forecasts (Figure 1a). To implement the pipeline outlined in Figure 1a, we used a "continuous analysis" framework [*sensu* @beaulieu2017] that automatically processes the most up-to-date data, refits the models, makes new forecasts, archives the forecasts, and updates a website with analysis of current and previous forecasts. In this section we describe our approach to streamlining and automating the multiple components of the forecasting pipeline and the tools and infrastructure we employed to execute each stage of the pipeline.

![Figure 1. a) Stages of the forecasting pipeline. To go from raw data to forecast presentation involves a number of stages, each of which requires unique tasks, tools and infrastructure. While each stage has unique tasks associated with it, the stages of our pipeline are also interdependent. Our pipeline is primarily a linear structure where outputs from one stage form the inputs for the subsequent stage. All stages require code written in R. b) Continuous integration system. Each box denotes the core infrastructure used for each stage of the forecasting pipeline denoted in Figure 1a. Continuous integration, denoted here with the Travis icon (a woman wearing safety glasses and hardhat), plugs into our pipeline at every stage.  Travis triggers the code involved in major events that link the stages of the pipeline, such as taking the output from the forecasting stage (purple box) to create an updated presentation (rose box). Travis also runs tasks within a stage, such as conducting tests to make sure code changes have not introduced errors (icons on arrows originating and ending on the same box). ](figs/Figure1.png)

@@

A core aspect of iterative near-term forecasting is the regular rerunning of the forecasting pipeline. We chose to have the computer run our pipeline and employed "continuous analysis" [*sensu* @beaulieu2017] to drive the automation of both the full pipeline and a number of its individual components. Continuous analysis uses a set of tools originally designed for software development called "continuous integration" (CI). CI combines computing environments for running code with monitoring systems to identify changes in data or code. Essentially, CI is a computer helper whose job is to watch the pipeline and, when it sees a change in the code or data, it runs all the computer scripts needed to ensure that the forecasting pipeline runs from beginning to end. This is useful for iterative near-term forecasting because it does not rely on humans to remember to create forecasts when new models or data are added. These tools are common in the area of software development where they are used to automate software testing and integrate work by multiple developers working on the same code base. However, these tools can be used for any computational task that needs to be regularly repeated or run after changes to code or data [@beaulieu2017]. Our forecasting pipeline currently runs on a publicly available continuous integration service (Travis CI; https://travis-ci.org/) that is free for open source projects (up to a limited amount of computing time). Because of the widespread use of CI in software development, alternative services that can run code on local or cloud-based computational infrastructure also exist (e.g., Drone (http://try.drone.io/)) [@beaulieu2017]. As detailed below, we use CI to quality check data, test code using "unit tests" [@wilson2014], build models, make forecasts, and publicly present and archive the results (Figure 1b).

@@

New data are double-entered into Microsoft Excel using the "data validation" feature. The two versions are then compared using an R script to control for errors in data entry. Quality control (QC) checks using the `testthat` R package [@wickham2011] are run on the data to test for validity and consistency both within the new data and between the new and archived data. The local use of the QC scripts to flag problematic data greatly reduces the time spent error-checking and ensures that the quality of data is consistent. The data are then uploaded to the GitHub-based Portal Data repository (https://github.com/weecology/PortalData). GitHub (https://github.com/) is a software development tool for managing computer code development, but we have also found it useful for data management. On GitHub, changes to data can be tracked through the Git version control system which logs all changes made to any files in the repository - giving us a record of exactly of when specific lines of data were changed or added. All updates to data are processed through "pull requests", which are notifications that someone has a modified version of the data to contribute. QA/QC checks are automatically rerun on the submitted data using continuous integration to ensure that these checks are run and that no avoidable errors reach the official version of the dataset.

@@

The Portal Project has a long history of making its data publicly available, so that anyone can use it for forecasting or other projects. Historically the publication of the data was conducted through data papers [@ernest2009,@ernest2016], the most common approach in ecology; however, this approach caused years of data latency. Recently, the project has switched to posting data directly to a public GitHub repository with a CC0 (i.e. no rights reserved) license (Figure 1). This immediate posting reduces that data latency to less than one month and, therefore, makes meaningful iterative near-term forecasting possible for not only our group but other interested parties, as well.

@@

Once data is available, it must be processed into a form appropriate for modeling (Figure 1). For many ecological datasets, this requires not only simple data manipulation but also a good understanding of the data to facilitate appropriate aggregation. Data manipulation steps are often conducted using custom one-off code to convert the raw data into the desired form [@morris2013], but this approach has several limitations. First, each researcher must develop and maintain their own data manipulation code, which is inefficient and can result in different researchers producing different versions of the data for the same task. Subtle differences in data processing decisions have led to confusion when reproducing results for the Portal data in the past. Second, this kind of code is rarely robust to changes in data structure and location. Based on our experience developing and maintaining the Data Retriever [@morris2013; @senyondo2017], these kinds of changes are common. Finally, this kind of code is generally poorly tested, which can lead to errors based on mistakes in data manipulation. To avoid these issues for the Portal Project data, the Portal team has been developing an R package (portalR; http://github.com/weecology/portalr) for acquiring the data and handling common data cleaning and aggregation tasks. As a result, our modeling and forecasting code only needs to install this package and run the data manipulation and summary functions to get the appropriate data (Figure 1b). The package undergoes thorough automated unit tests to ensure that data manipulations are achieving the desired results. Having data manipulation code maintained in a separate package that focuses on consistently providing properly summarized forms of the most recent data has made maintaining the forecasting code itself much more straightforward.

@@

Iterative near-term forecasting involves regularly refitting a variety of different models (Figure 1). Ideally, new models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach. We use CI to refit the models and make new forecasts each time the modeling code changes and when new data become available (Figure 1b). We use a plugin infrastructure to allow new models to be easily added to the system. This approach treats each model as an interchangable black box - all models have access to the same input data and generate the same structure for model outputs (Figure 2). Each iteration of the forecasting cycle, all existing models are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts. A weighted ensemble model is then added with weights based on how well individual models fit the training data. This plugin infrastructure makes it easy to add and compare very different types of models, from the basic time-series approaches currently implemented to the more complex state-space and machine learning models we hope to implement in the future. As long as a model script can load the provided data and produce the appropriate output it will be run and its results incorporated into the rest of the forecasting system.

![Figure 2. Demonstration of plugin infrastructure. All model scripts (represented here by the example AR1.R) are housed in a single folder. Each model script uses data provided by the core forecasting code (represented here by rodent.csv) and returns its forecast outputs in a predefined structure that is consistent across models (represented here by the example 2017_12_08forecasts.csv). Outputs from all models run on a particular date are combined into the same file (i.e. 2017_12_08forecasts.csv) to allow cross-model evaluations. Model output files are housed in a folder containing all forecast outputs from all previous dates to facilitate our arhciving process and forecast assessment activities ](figs/plugin_fig.gif)

In addition to flexibility in what model structures can be supported, we also wanted to support flexibility in what the models predict. Allowing models to make forecasts for system properties ranging from individual species' population abundances to total community biomass facilitates exploration of differences in forecastability across different aspects of ecological systems. We designed a forecast output format to support this. Each forecast output file contains the date being forecast, the collection date of the data used for fitting the models, the date the forecast was made, the state variable being forecast (e.g., rodent biomass, the abundance of a species), and the forecast value and associated uncertainty of that forecast (Figure 2). This allows us to store a variety of different forecasts in a common format and may serve as a useful starting point for developing a standard for storing ecological forecasts more generally.

Forecasts are currently evaluated using root mean square error (RMSE) to evaluate point forecasts and coverage to evaluate uncertainty. We plan to add additional metrics in the future. In addition to evaluating the actual forecasts, we also use hindcasting (forecasting on already collected data) to gain additional insight into the methods that work best for forecasting this system. For example, a model is fit using rodent observations up to June 2005, then used to make a forecast 12 months out to May 2006. The observations of that 12 month period can immediately be used to evaluate the model. Hindcasting can be conducted using data that was already been collected, thus allowing model comparison of large numbers of hindcasts and giving insight into which models make the best forecasts without needing to wait for new data to be collected. It can also be used to quickly evaluate new models instead of waiting for an adequate amount of data to accumulate.

@@

Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2018; @harris2018] (Figure 1). Archiving serves as a form of pre-registration for model predictions, because the forecasts cannot be modified once the data to assess them has been collected. This helps facilitate an unbiased interpretation of model performance. To serve this role archives should be publicly accessible and be a permanent record that cannot be changed or deleted. The second criteria means that GitHub is not sufficient for archival purposes because repositories can changed or deleted [@bergman2012].  We explored three major repositories for archiving forecasts: FigShare (https://figshare.com/), Zenodo (https://zenodo.org/), and Open Science Framework (https://osf.io/). While all three repositories allowed for easy manual submissions (i.e., a human uploading files after each forecast), automating this process was substantially more difficult. Various combinations of repositories, APIs (i.e., interfaces for automatically interacting with these websites) and associated R packages had issues with: 1) integrating authorization with continuous integration; 2) automatically making archived files public; 3) adding new files to an existing location; or 4) automatically permanently archiving the files. Our eventual solution was to leverage the GitHub-Zenodo integration (https://guides.github.com/activities/citable-code/) and automatically push forecasts to a GitHub repository from the CI server and release them via the GitHub API. The GitHub-Zenodo integration is designed to automatically create versioned archives of GitHub repositories. We created a repository for storing forecasts (https://github.com/weecology/forecasts) and linked this repository with Zenodo (a one-time manual process). Each time a new forecast is created, our pipeline adds the new forecasts to the GitHub repository and uses the GitHub API to create a new "release" for that repository. This triggers the Zenodo-GitHub integration, which automatically archives the resulting forecasts under a top-level DOI that refers to all archived forecasts ([https://doi.org/10.5281/zenodo.839580](https://doi.org/10.5281/zenodo.839580)). Through this process, we automatically archive every forecast made with a documented time-stamp for every forecast in the archive. In addition, we also archive the full state of the repository for modeling and forecasting ([https://doi.org/10.5281/zenodo.833438](https://doi.org/10.5281/zenodo.833438)). This ensures that every forecast is fully reproducible since the exact code used to generate every forecast is preserved. Early forecasts from this system are archived as part of this modeling and forecasting code archive.

@@

Each month, we present our forecasts on a website that displays monthly rodent forecasts, model evaluation metrics, monthly reports, and information about the study site (Figure 3; [http://portal.naturecast.org](http://portal.naturecast.org/)). The website includes a graphical presentation of the most recent month's forecasts (including uncertainty) and compares the latest data to the previous forecasts. Information on the species and the field site targeted to a general audience are also included. The site is built using Rmarkdown [@rmarkdown], which naturally integrates into the pipeline, and is automatically updated after each forecast. The `knitr` R package [@xie2015] compiles the code into HTML, which is then published using Github Pages (https://pages.github.com/). The files for the website are stored in a subdirectory of the forecasting repository. As a result, the website is also archived automatically as part of archiving the forecast results.

@@

Following the recommendations of Dietze et al [-@dietze2018], we developed an automated iterative forecasting system (Figure 1) to support repeated forecasting of an ecological system. Our forecasting system automatically acquires and processes the newest data, refits the models, makes new forecasts, publicly archives those forecasts, and presents both the current forecast and information on how previous forecasts performed. Every week our forecasting system generates a new set of forecasts with no human intervention, except for the entry of new field data. Our forecasting system ensures that forecasts based on the most recent data are always available.  Our forecasting system is also designed to allow rapid assessment of the performance of multiple forecasting models for a number of different states of the system, including the abundances of individual species and community-level variables such as total abundance. To create this iterative near-term forecasting system, we used R to process data and conduct analyses, and we leveraged already existing services (i.e. GitHub, Travis, Docker) for more complicated cyberinfrastructure tasks. Thus, our approach to developing iterative near-term forecasting infrastructure provides an example for how short-term ecological forecasting systems can be initially developed.

>>>>

---
title: "Automated iterative near-term forecasting for the Portal Project"
csl: mee.csl
fontsize: 12pt
output:
  pdf_document:
    keep_tex: yes
  html_document: default
header-includes:
- \usepackage{times}
- \usepackage{setspace}
- \usepackage{booktabs}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
bibliography: refs.bib
---
\raggedright

@@

Forecasting the future state of ecological systems is important for management, conservation, and evaluation of how well models capture the processes governing ecological systems [@clark2001; @tallis2006; @diaz2015; @dietze2017]. In 2001, Clark et al. [-@clark2001] called for a more central role of forecasting in ecology, and since then an increasing number of ecological forecasts are being published. However, most of these forecasts are made once, published, and never assessed or updated. This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions. The lack of active assessment results in limited information on how much confidence to place in forecasts and makes if difficult to determine which forecasting methods to build on. Without regular updates, forecasts lack the most current data and forecast accurracy will decay over time [@petch2015; @dietze2018]. More regular updating and assessment will advance ecological forecasting as a field by accelerating the identification of the best models for individual forecasts and improving our understanding of how to best design forecasting approaches for ecology in general. For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].

With the goal of making ecological forecasting more dynamic and responsive, Dietze et al [-@dietze2018] recently called for an increase in iterative near-term forecasting. Iterative near-term forecasting is defined by making predictions for the near future and repeatedly updating those predictions through a cycle of evaluation, integration of new data, and generation of new forecasts. Because forecasts are made 'near-term'---daily to annual time scales instead of multi-decadal---predictions can be assessed more quickly and frequently, leading to more rapid model improvements [@tredennick2016; @dietze2018]. In addition, since forecasts are made repeatedly through time, new data can be continuosly integrated with each iteration [@dietze2018]. As a result, iterative near-term forecasting has the potential to promote rapid improvement in the state of ecological forecasting by quickly identifying how models are failing, facilitating rapid testing of improved models, and incorporating the most up-to-date data available. In addition to yielding improved information for guiding policy and management [@clark2001; @luo2011; @petchey2015], this iterative approach will help improve our basic understanding of ecological systems [@dietze2018]. For example, alternative mechanistic models can be compared to determine which model provides the best forecasts, thus providing insights into the importance of different ecological processes [@dietze2018]. Iterative near-term forecasting provides a more dynamic interplay between models, predictions, and data that will improve ecological forecasting and our understanding of ecological systems more broadly.

Because iterative near-term forecasting requires a dynamic integration of models, predictions, and data, Dietze et al [-@dietze2018] highlight approaches to data management, model construction and evaluation, and cyberinfrastructure that are necessary to effectively implement this type of forecasting (Box 1). Data needs to be released quickly under open licenses [@vargas2017; @dietze2018] and structured so that it can be used easily by a variety of researchers and in multiple modeling approaches [@borer2009; @strasser2011]. Models need to be able to deal with uncertainty, in the predictors and the predictions, to properly convey uncertainty in the resulting forecasts [@diniz-filho2009]. Multiple models should be developed, both to assess which models are performing best [@dietze2018] and to facilitate combining models to form ensemble predictions which tend to perform better than single models [@araujo2007; @diniz-filho2009]. Ensuring that data and models are regularly updated and new forecasts are made requires cyberinfrastructure to automate data processing, model fitting, prediction, model evaluation, forecast visualization, and archiving. In combination, these approaches should allow forecasts to be easily rerun and evaluated as new data becomes available [Box 1; @dietze2018].

While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement and few of their recommendations are in widespread use in ecology today. We explored what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going long-term ecological study that collects high-frequency data on desert rodent abundances [@brown1998; @ernest2008]. We constructed an automated forecasting pipeline with the goal of being able to forecast rodent abundances and evaluate our predictions on a monthly basis. In this paper, we discuss our approach for creating this iterative near-term forecasting pipeline, the challenges we encountered, the tools we used, and the lessons we learned so that others can create their own iterative forecasting systems.

@@

Iterative forecasting is most effective with frequently collected data, since it provides more opportunities for updating model results and assessing (and potentially improving) model performance [Box 1; @dietze2018]. The Portal Project is a long-term ecological study situated in the Chihuahuan Desert (2 km north and 6.5 km east of Portal, Arizona, US). Researchers have been continuously collecting data at the site since 1977, including data on the abundance of rodent and plant species (monthly and twice yearly, respectively) and climatic factors such as air temperature and precipitation (daily) [@brown1998; @ernest2009; @ernest2016]. The site consists of 24 50m x 50m experimental plots. Each plot contains 49 permanently marked trapping stations laid out in a 7 x 7 grid, and all plots are trapped with Sherman live traps for one night each month. For all rodents caught during a trapping session, information on species identity, size, and reproductive condition is collected, and new individuals are given identification tags. This information on rodent populations is high-frequency, uses consistent trapping methodology, and has an extended time-series (469 monthly samples and counting), making this study an ideal case for near-term iterative forecasting.

@@

Implementation of iterative forecasting requires the regular rebuilding of models with new raw data as it becomes available and the presentation of those forecasts in usable forms; in our case, this occurs monthly. Rebuilding models in an efficient and maintainable way relies on developing an automated pipeline to handle the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presention of the forecasts (Figure 1a). To implement the pipeline outlined in Figure 1a, we used a "continuous analysis" framework [*sensu* @beaulieu2017] that automatically processes the most up-to-date data, refits the models, makes new forecasts, archives the forecasts, and updates a website with analysis of current and previous forecasts. In this section we describe our approach to streamlining and automating the multiple components of the forecasting pipeline and the tools and infrastructure we employed to execute each component.

![Figure 1. a) Stages of the forecasting pipeline. To go from raw data to forecast presentation involves a number of stages, each of which requires unique tasks, tools and infrastructure. The stages of are interdependent with outputs from one stage forming the inputs for the subsequent stage. Tasks in all stages are run using code written in R. b) Continuous integration system. Each box denotes the core infrastructure used for each stage of the forecasting pipeline. Continuous integration (denoted by the Travis icon; a woman wearing safety glasses and hardhat) triggers the code involved in events that link the stages of the pipeline, such as using the output from the forecasting stage (purple box) to create an updated website (rose box). Travis also runs tasks within a stage, such as testing code and adding weather data (icons on arrows originating and ending on the same box). ](figs/Figure1.png)

@@

A core aspect of iterative near-term forecasting is the regular rerunning of the forecasting pipeline. We employed "continuous analysis" [*sensu* @beaulieu2017] to drive the automation of both the full pipeline and a number of its individual components. Continuous analysis uses a set of tools originally designed for software development called "continuous integration" (CI). CI combines computing environments for running code with monitoring systems to identify changes in data or code. Essentially, CI is a computer helper whose job is to watch the pipeline and, when it sees a change in the code or data, it runs all the computer scripts needed to ensure that the forecasting pipeline runs from beginning to end. This is useful for iterative near-term forecasting because it does not rely on humans to create new forecasts whenever new models or data are added. These tools are common in the area of software development where they are used to automate software testing and integrate work by multiple developers working on the same code base. However, these tools can be used for any computational task that needs to be regularly repeated or run after changes to code or data [@beaulieu2017]. Our forecasting pipeline currently runs on a publicly available continuous integration service (Travis CI; https://travis-ci.org/) that is free for open source projects (up to a limited amount of computing time). Because of the widespread use of CI in software development, alternative services that can run code on local or cloud-based computational infrastructure also exist [@beaulieu2017]. We use CI to quality check data, test code using "unit tests" [@wilson2014], build models, make forecasts, and publicly present and archive the results (Figure 1b).

@@

New data are double-entered into Microsoft Excel using the "data validation" feature. The two versions are then compared using an R script to control for errors in data entry. Quality control (QC) checks using the `testthat` R package [@wickham2011] are run on the data to test for validity and consistency both within the new data and between the new and archived data. The local use of the QC scripts to flag problematic data greatly reduces the time spent error-checking and ensures that the quality of data is consistent. The data are then uploaded to the GitHub-based Portal Data repository (https://github.com/weecology/PortalData). GitHub (https://github.com/) is a software development tool for managing computer code development, but we have also found it useful for data management. On GitHub, changes to data can be tracked through the Git version control system which logs all changes made to any files in the repository - giving us a record of exactly of when specific lines of data were changed or added. All updates to data are processed through "pull requests", which are notifications that someone has a modified version of the data to contribute. QA/QC checks are automatically run on the submitted data using continuous integration to ensure that these checks are run and that no avoidable errors reach the official version of the dataset.

@@

The Portal Project has a long history of making its data publicly available, so that anyone can use it for forecasting or other projects. Historically the publication of the data was conducted through data papers [@ernest2009,@ernest2016], the most common approach in ecology; however, this approach caused years of data latency. With the recent switch to posting data directly to a public GitHub repository (Figure 1) with a CC0 waiver (i.e. no restrictions on data use; https://creativecommons.org/publicdomain/zero/1.0/) data latency for everyone has been reduced to less than one month making meaningful iterative near-term forecasting possible for not only our group but other interested parties, as well.

@@

Once data is available, it must be processed into a form appropriate for modeling (Figure 1). For many ecological datasets, this requires not only simple data manipulation but also a good understanding of the data to facilitate appropriate aggregation. Data manipulation steps are often conducted using custom one-off code to convert the raw data into the desired form [@morris2013], but this approach has several limitations. First, each researcher must develop and maintain their own data manipulation code, which is inefficient and can result in different researchers producing different versions of the data for the same task. Subtle differences in data processing decisions have led to confusion when reproducing results for the Portal data in the past. Second, this kind of code is rarely robust to changes in data structure and location. Based on our experience developing and maintaining the Data Retriever [@morris2013; @senyondo2017], these kinds of changes are common. Finally, this kind of code is generally poorly tested, which can lead to errors based on mistakes in data manipulation. To avoid these issues for the Portal Project data, the Portal team has been developing an R package (portalr; http://github.com/weecology/portalr) for acquiring the data and handling common data cleaning and aggregation tasks. As a result, our modeling and forecasting code only needs to install this package and run the data manipulation and summary functions to get the appropriate data (Figure 1b). The package undergoes thorough automated unit testing to ensure that data manipulations are achieving the desired results. Having data manipulation code maintained in a separate package that focuses on consistently providing properly summarized forms of the most recent data has made maintaining the forecasting code itself much more straightforward.

@@

Iterative near-term forecasting involves regularly refitting a variety of different models (Figure 1). Ideally, new models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach. We use CI to refit the models and make new forecasts each time the modeling code changes and when new data become available (Figure 1b). We use a plugin infrastructure to allow new models to be easily added to the system. This approach treats each model as an interchangable black box - all models have access to the same input data and generate the same structure for model outputs (Figure 2). During each run of the forecasting code, all existing models are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts. A weighted ensemble model is then added with weights based on how well individual models fit the training data. This plugin infrastructure makes it easy to add and compare very different types of models, from the basic time-series approaches currently implemented to the more complex state-space and machine learning models we hope to implement in the future. As long as a model script can load the provided data and produce the appropriate output it will be run and its results incorporated into the rest of the forecasting system.

![Figure 2. Demonstration of plugin infrastructure. All model scripts (represented here by the example AR1.R) are housed in a single folder. Each model script uses data provided by the core forecasting code (represented here by rodent.csv) and returns its forecast outputs in a predefined structure that is consistent across models (represented here by the example 2017_12_08forecasts.csv). Outputs from all models run on a particular date are combined into the same file (i.e. 2017_12_08forecasts.csv) to allow cross-model evaluations. Model output files are housed in a folder containing all forecast outputs from all previous dates to facilitate archiving and forecast assessment.](figs/plugin_fig.gif)

In addition to flexibility in what model structures can be supported, we also wanted to support flexibility in what the models predict. Allowing models to make forecasts for system properties ranging from individual species' population abundances to total community biomass facilitates exploration of differences in forecastability across different aspects of ecological systems. We designed a forecast output format to support this. Each forecast output file contains the date being forecast, the collection date of the data used for fitting the models, the model name, the date the forecast was made, the state variable being forecast (e.g., rodent biomass, the abundance of a species), and the forecast value and associated uncertainty of that forecast (Figure 2). This allows us to store a variety of different forecasts in a common format and may serve as a useful starting point for developing a standard for storing ecological forecasts more generally.

Forecasts are currently evaluated using root mean square error (RMSE) to evaluate point forecasts and coverage to evaluate uncertainty. We plan to add additional metrics in the future. In addition to evaluating the actual forecasts, we also use hindcasting [forecasting on already collected data; @jolliffe2003] to gain additional insight into the methods that work best for forecasting this system. For example, a model is fit using rodent observations up to June 2005, then used to make a forecast 12 months out to May 2006. The observations of that 12 month period can immediately be used to evaluate the model. Since hindcasting is conducted using data that has already been collected, it allows model comparisons to be conducted on large numbers of hindcasts and provides insight into which models make the best forecasts without needing to wait for new data to be collected [@harris2018]. It can also be used to quickly evaluate new models instead of waiting for an adequate amount of data to accumulate.

@@

Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2018; @harris2018] (Figure 1). Archiving serves as a form of pre-registration for model predictions, because the forecasts cannot be modified once the data to assess them has been collected. This helps facilitate an unbiased interpretation of model performance. To serve this role archives should be publicly accessible and be a permanent record that cannot be changed or deleted. The second criteria means that GitHub is not sufficient for archival purposes because repositories can be changed or deleted [@bergman2012; @white2015].  We explored three major repositories for archiving forecasts: FigShare (https://figshare.com/), Zenodo (https://zenodo.org/), and Open Science Framework (https://osf.io/). While all three repositories allowed for easy manual submissions (i.e., a human uploading files after each forecast), automating this process was substantially more difficult. Various combinations of repositories, APIs (i.e., interfaces for automatically interacting with the archiving websites) and associated R packages had issues with: 1) integrating authorization with continuous integration; 2) automatically making archived files public; 3) adding new files to an existing location; or 4) automatically permanently archiving the files. Our eventual solution was to leverage the GitHub-Zenodo integration (https://guides.github.com/activities/citable-code/) and automatically push forecasts to a GitHub repository from the CI server and release them via the GitHub API. The GitHub-Zenodo integration is designed to automatically create versioned archives of GitHub repositories. We created a repository for storing forecasts (https://github.com/weecology/forecasts) and linked this repository with Zenodo (a one-time manual process). Each time a new forecast is created, our pipeline adds the new forecasts to the GitHub repository and uses the GitHub API to create a new "release" for that repository. This triggers the Zenodo-GitHub integration, which automatically archives the resulting forecasts under a top-level DOI that refers to all archived forecasts ([https://doi.org/10.5281/zenodo.839580](https://doi.org/10.5281/zenodo.839580)). Through this process, we automatically archive every forecast made with a documented time-stamp. In addition, we also archive the full state of the modeling and forecasting repository ([https://doi.org/10.5281/zenodo.833438](https://doi.org/10.5281/zenodo.833438)). This ensures that every forecast is fully reproducible since the exact code used to generate every forecast is preserved. Early forecasts from this system are archived in this modeling and forecasting code archive, not the newer forecasts repository.

@@

Each month, we present our forecasts on a website that displays monthly rodent forecasts, model evaluation metrics, monthly reports, and information about the study site (Figure 3; [http://portal.naturecast.org](http://portal.naturecast.org/)). The website includes a graphical presentation of the most recent month's forecasts (including uncertainty) and compares the latest data to the previous forecasts. Information on the species and the field site are also included. The site is built using Rmarkdown [@rmarkdown], which naturally integrates into the pipeline, and is automatically updated after each forecast. The `knitr` R package [@xie2015] compiles the code into HTML, which is then published using Github Pages (https://pages.github.com/). The files for the website are stored in a subdirectory of the forecasting repository. As a result, the website is also archived automatically as part of archiving the forecast results.

@@

Following the recommendations of Dietze et al [-@dietze2018], we developed an automated iterative forecasting system (Figure 1) to support repeated forecasting of an ecological system. Our forecasting system automatically acquires and processes the newest data, refits the models, makes new forecasts, publicly archives those forecasts, and presents both the current forecast and information on how previous forecasts performed. Every week our forecasting system generates a new set of forecasts with no human intervention, except for the entry of new field data. Our forecasting system ensures that forecasts based on the most recent data are always available.  It is also designed to allow rapid assessment of the performance of multiple forecasting models for a number of different states of the system, including the abundances of individual species and community-level variables such as total abundance. To create this iterative near-term forecasting system, we used R to process data and conduct analyses, and leveraged existing tools and services (i.e. GitHub, Travis, Docker) for more complicated cyberinfrastructure tasks. Thus, our approach to developing iterative near-term forecasting infrastructure provides an example for how short-term ecological forecasting systems can be developed.
